---
output: 
    github_document:
      html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# onpoint

<!-- badges: start -->

README Last updated: `r Sys.Date()`

[![R-CMD-check](https://github.com/r-spatialecology/onpoint/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/r-spatialecology/onpoint/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/r-spatialecology/onpoint/branch/main/graph/badge.svg?token=RkgCfHnPyf)](https://codecov.io/gh/r-spatialecology/onpoint)
[![Project Status](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![lifecycle](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![CRAN status](https://www.r-pkg.org/badges/version/onpoint)](https://CRAN.R-project.org/package=onpoint)
[![CRAN logs](https://cranlogs.r-pkg.org/badges/grand-total/onpoint)](http://cran.rstudio.com/web/packages/onpoint/index.html)
[![License](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)

<!-- badges: end -->

`onpoint` is a growing collection of helper functions mainly related to point pattern analysis and the  [spatstat](http://spatstat.org/) package.

## Installation

You can install the released version of `onpoint` from CRAN using:

```{r installation_cran, eval = FALSE}
install.packages("onpoint")
```

You can install the development version of `onpoint` from [Github](https://github.com/r-spatialecology/onpoint) using:

```{r installation_github, eval = FALSE}
# install.packages("remotes")
remotes::install_github("r-spatialecology/onpoint")
```

```{r, message = FALSE}
library(onpoint)
library(spatstat)

data(spruces)
```

### Summary functions 

Currently, `onpoint` provides three second-order summary functions, namely Besag's L-function centered to zero and the O-ring statistic and a fast estimation of the pair-correlation function. 

Centering Besag's L-function to zero has the advantage of an easier interpretation and plotting (Haase 1995). The function `center_Lest()` can either deal with a point pattern and calculated the centered L-function directly, or center the L-function afterwards it was calculating using `spatstat`s `Lest()`.

```{r calc_lfun}
# calculate L-function
l_function <- Lest(spruces, correction = "Ripley")

# center L-function to zero
l_function_centered <- center_Lest(spruces, correction = "Ripley")
```

```{r plot_lfun, echo = FALSE, fig.width = 10, fig.height = 5, fig.align = "center"}
par(mfrow = c(1, 2))
plot(l_function, main = "Besag's L-function")
plot(l_function_centered, main = "centered Besag's L-function")
par(mfrow = c(1, 1))
```

The O-ring statistic O(r) (Wiegand & Moloney 2004) can be calculated using `Oest()`. Generally speaking, O(r) scales the pair correlation g(r) function with help of the intensity \eqn{\lambda}. One advantage of the O-ring statistic is that it can be interpreted as a neighborhood density because it is a probability density function (Wiegand & Moloney 2004).

```{r calc_oring}
o_ring <- Oest(spruces)
```

Of course, both summary functions can be used in combination with `spatstat`'s `envelope()` function.

```{r env_oring}
oring_envelope <- envelope(spruces, fun = estimate_o_ring, nsim = 199, verbose = FALSE)
```

```{r plot_oring, echo = FALSE, fig.width = 7.5, fig.height = 5, fig.align = "center"}
plot(oring_envelope, main = "O-ring statistic")
```

`pcf_fast()` estimates the pair-correlation function based on Ripley's K-function, which is faster than estimation the pair-correlation function directly.

```{r calc_fast, collapse = TRUE}
pcf_fast(spruces)
```

### Null models 

`onpoint` includes two functions to simulate null model patterns. 

`rheteroppp()` is a convienent wrapper around a few `spatstat` functions to straighforward simulate a heterogeneous Poisson process.

```{r calc_hetero}
null_model_hetero <- rheteroppp(spruces, nsim = 199)

hetero <- envelope(spruces, fun = pcf, 
                   funargs = list(correction = "Ripley", divisor = "d"),
                   simulate = null_model_hetero, nsim = 199, 
                   verbose = FALSE)
```

```{r plot_hetero, echo = FALSE, fig.width = 7.5, fig.height = 5, fig.align = "center"}
plot(hetero, main = "Heterogeneous Poisson process")
```

To simulate antecedent conditions in which only one pattern influences the other, but not the other way around (Wiegand & Moloney 2004, Velazquez et al. 2016), `simulate_antecedent()` can be used. This null model randomizes only one type of points (e.g. seedlings), while keeping the other type of points constant (e.g. mature trees) to check for associations between the two.

```{r calc_antecedent}
marks(spruces) <- ifelse(marks(spruces) > 0.3, yes = "adult", no = "seedling")

null_model_antecedent <- simulate_antecedent(spruces, i = "seedling", j = "adult",
                                             nsim = 199)

antecedent <- envelope(spruces, fun = pcf, 
                       funargs = list(correction = "Ripley", divisor = "d"),
                       simulate = null_model_antecedent, nsim = 199, 
                       verbose = FALSE)
```

```{r plot_antecedent, echo = FALSE, fig.width = 7.5, fig.height = 5, fig.align = "center"}
plot(antecedent, main = "Antecedent conditions")
```

### Various

To plot simulation envelopes using quantum plots (e.g., Esser et al. 2015), just pass an `envelope` object as input to `plot_quantums()`.

```{r plot_quantums, fig.width = 7.5, fig.height = 5, fig.align = "center"}
plot_quantums(antecedent, ylab = "g(r)")
```

### Contributing

Please note that the **onpoint** package is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.

To contribute to this project, please see the [Contributing guidelines](CONTRIBUTING.md).

#### References
Besag, J.E., 1977. Discussion on Dr. Ripley’s paper. Journal of the Royal Statistical Society. Series B (Methodological) 39, 193–195. <https://doi.org/10.1111/j.2517-6161.1977.tb01616.x>

Esser, D.S., Leveau, J.H.J., Meyer, K.M., Wiegand, K., 2015. Spatial scales of interactions among bacteria and between bacteria and the leaf surface. FEMS Microbiology Ecology 91, 1–13. <https://doi.org/10.1093/femsec/fiu034>

Haase, P., 1995. Spatial pattern analysis in ecology based on Ripley’s K-function: Introduction and methods of edge correction. Journal of Vegetation Science 6, 575–582. <https://doi.org/10.2307/3236356>

Velázquez, E., Martínez, I., Getzin, S., Moloney, K.A., Wiegand, T., 2016. An evaluation of the state of spatial point pattern analysis in ecology. Ecography 39, 1–14. <https://doi.org/10.1111/ecog.01579>

Wiegand, T., Moloney, K.A., 2004. Rings, circles, and null models for point pattern analysis in ecology. Oikos 104, 209–229. <https://doi.org/10.1111/j.0030-1299.2004.12497.x>

